Pulsar Acceleration Shifts from nearby Supernova Explosion

### 2016-02-05 draft3_1C.tex

With the forecast of thousands of millisecond pulsars to be found by the Square Kilometer Array \cite{SKA}, a new SN effectively sits within a timing-grid. At the instant the SNe is detected at earth, even if only from direct neutrino detection, we know that all pulsars signals enter the first stage of changes, by different amounts depending on their impact parameter. The closest or the best timed pulsars may fit our result and determine $\delta M$ within a decade or two. The sharp change in the derivative of period will also happen around the same time scale for those with small $\theta$. Given a few SNe per century \cite{SNrate06}, a signature detection of this two-consecutive-acceleration-shifts is possible in the foreseeable future.

>>>>

The Square Kilometer Array \cite{SKA} and its future extension \cite{MSPpopulation} forecast $\sim 30000$ millisecond pulsars (MSP) throughout our galaxy. They mostly distribute in the galactic center, a region with height $\sim 500$pc and radius $\sim 7.5$kpc. With a typical separation $\sim 200$ light years from each other, these MSPs form a timing grid. Since we also expect the next SN to be likely near the galactic center, with such a dense timing grid, it is almost guaranteed that the SN will be behind some MSP with $(r_p\sin\theta) < 100$ light years, which already makes the first stage of timing shift quite detectible \cite{OluPie13}. Actually, since we expect both the chance of SNe and the density of MSPs to be even higher near the very center \footnote{The forecast simulation \cite{MSPpopulation} actually cut-off the very center region.}, it is not unreasonable to expect $r_p\sim30$ light years, which makes the second stage change observable.

This means that as soon as we see the next SNe, we can start to follow the first stage timing shift in all MSPs. Some of them will show the predicted behavior within a few decades. By about the same time, likely one or two among those will go through the second stage, having a sudden change in the second derivative of timing shifts. Given a few SNe per century \cite{SNrate06}, such a signature detection is in the foreseeable future.
